---
title: "rmd test"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


```{r, eval=TRUE}
library('readr')
library('ggplot2')
library('magrittr')
library('tidyverse')
library('lubridate')
library('dplyr')

```

Import data
```{r, eval= TRUE}
landings_data <- read_csv("sample_landings_data_raw.csv")

landings_data
```

Start with the landings_data data frame and rename columns. Also turn the date column into a date format that R recognises.
```{r, eval= TRUE}
landings_data <- landings_data %>%
 rename( Year = yy,
         Date = dat,
         Trip_ID = trip,
         Effort_Hours = effort,
         Gear = gr,
         Species = sp,
         Length_cm = l_cm,
         Weight_g = w_cm) %>%
  mutate(Date = mdy(Date))
```
Print data.
```{r}
landings_data
```
Check for missing values
```{r}
landings_data[!complete.cases(landings_data),]
```
Remove rows with missing values
```{r}
landings_data <- na.omit(landings_data)
```
Print edited data set
```{r}
landings_data
```

Check for typos in the gear variable
```{r}
unique(landings_data$Gear)
```
Trap appears twice - fix this
```{r}
landings_data <- landings_data %>%
  mutate(Gear = tolower(Gear))
unique(landings_data$Gear)
```
Check for typos in species variable
```{r}
unique(landings_data$Species)
```
Check how many times each spelling occurs
```{r}
landings_data %>%
  filter(Species == "Caesoi cunning") %>%
  nrow()
landings_data %>%
  filter(Species == "Caesio cuning") %>%
  nrow()
```
Caesio cuning is typo as it appears only twice, replace with correct spelling 
```{r}
landings_data <- landings_data %>%
  mutate(Species = replace(Species,Species == "Caesoi cunning", "Caesio cuning"))

unique(landings_data$Species)
```

Look at the distribution of length
```{r}
summary(landings_data$Length_cm)
```
The max value is much larger than the mean/median values. Make a plot to check for outliers.
```{r}
plot(landings_data$Length_cm)
```

We see there is an outlier. Remove this error and make a new plot.
```{r}
landings_data <- landings_data %>%
  filter(Length_cm < 100)

plot(landings_data$Length_cm)
```

Examine data frame again
```{r}
landings_data
```

Save clean data
```{r}
write_csv(landings_data,"sample_landings_data_clean.csv")
```

